---
title: "Assignment 2: Task 2"
author: "Sophia Leiker"
date: "2/1/2022"
output:   
  html_document: 
    toc: yes
    code_folding: hide
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(patchwork)
library(cowplot)
library(GGally)
library(broom)
library(jtools)
library(caret)
library(AICcmodavg)
library(ggbeeswarm)
```

## 1. Introduction

This report will employ non linear least squares to estimate parameters of a length to weight model for lizard populations in New Mexico. To do this the  snout length to weight model will be used. 


**Data Source:** The data used comes from: [Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. ](https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f) and metadata, including raw data and methods used for data collection can be found here [EDI Data Portal](https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-jrn.210007001.37).

This data package contains data on lizards sampled by pitfall traps located at 11 consumer plots at Jornada Basin LTER site from 1989-2006. Lizards were captured in pitfall traps at the 11 LTER II/III consumer plots (a subset of NPP plots) quarterly for 2 weeks per quarter. Variables measured include species, sex, recapture status, snout-vent length, total length, weight, and whether tail is broken or whole.



```{r}
# Reading in the data
palmetto <- read_csv(here("data", "lizard.csv")) %>% 
  clean_names()

```

***

## 1. Fitting Snout Length to Weight Model

### A. Strategies for Initial guesses 

### B. Log Transform of data (OLS regression)

### C. Supplying NLS Start list with regression coefficients

***

## 2. Final Model: Femal and Male Spcies


***

## 3. Filtering for male Western Whiptail lizard (Cnemidophorus tigrisatus) - Fit NLS Model



## X. Data Citation

Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f

