---
title: "Assignment 2: Task 1"
author: "Sophia Leiker"
date: "1/30/2022"
output: 
  html_document: 
    toc: yes
    code_folding: hide
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(patchwork)
library(cowplot)
library(GGally)
library(broom)
library(jtools)
library(caret)
library(AICcmodavg)
```


## 1. Introduction

This report will employ binary logistic regression to test feasibility of using variables plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs) to classify whether a palmetto is species Serenoa repens or Sabal etonia.

The analysis will be performed twice, and will use cross validation to compare two models:

- Log odds of plant type using plant height, canopy length, canopy width and green leaves as predictor variable.
- Log odds of plant type using plant height, canopy width and green leaves (i.e., drop canopy length for this model)


**Data Source:** The data used comes from: [Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative](https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5) and metadata, including methods used for data collection can be found here [EDI Data Portal](https://portal.edirepository.org/nis/metadataviewer?packageid=edi.317.1).
The first dataset, palmetto_data, contains survival and growth data across multiple years, habitats and experimental treatments. The second dataset, seedlings_data, follows the fate of marked putative palmetto seedlings in the field to assess survivorship and growth. The final dataset, harvested_palmetto_data, contains size data and estimated dry mass (biomass in grams) of 33 destructively harvested palmetto plants (17 S. repens and 16 S. etonia) of varying sizes and across habitats.



```{r}
# Reading in the data
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  clean_names()

```

***


## 2. Data Exploration and Visualization

- Exploring differences in height, canopy length, canopy width, and green leaves for the two species
- `height` = maximum height 1981-2017 [cm]
- `length` = widest length of the canopy 1981-2017 [cm]
- `width` = widest width of the canopy perpendicular to the canopy length 1981-2017 [cm]
- `green_lvs` = count of green leaves 1981-2017 [dimensionless]
- Assessing what predictor variables might be best used for classification
- `species` column: 1 = Serenoa repens, 2 = Sabal etonia

### A. Cleaning up the data
- Rename to easily identify Serenoa repens to Sabal etonia
- Selecting only the columns of interest 

```{r}
palmetto_clean <- palmetto %>% 
  select("species", "height", "length", "width", "green_lvs")

palmetto_clean$species[palmetto_clean$species == 1] <- "repens"
palmetto_clean$species[palmetto_clean$species == 2] <- "etonia"


```

```{r eval=FALSE}
#Evaluating via ggpairs, eval = FALSE after visually assessing trends
palmetto_clean %>% 
  select(species, height:green_lvs) %>% 
  ggpairs(aes(color=species))

#seems like there are differences in green_lvs
#slight difference in height and length
#not much difference in width

```

```{r}
ggplot(data = palmetto_clean, aes(x = length, y = width)) +
  geom_point(aes(color = species)) +
  theme_minimal() +
  labs(x = "Length (cm)", y = "Width (cm)", title = "Length and Width for Serenoa repens vs Sabal etonia", subtitle = "Data from: Abrahamson, W.G. 2019")
```

**Figure 1**: ADD TEXT HERE

```{r}
ggplot(data = palmetto_clean, aes(x = height, y = green_lvs)) +
  geom_point(aes(color = species)) +
  theme_minimal() +
  labs(x = "Height (cm)", y = "Green Leaves (count)", title = "Height and Green Leaves Count for Serenoa repens vs Sabal etonia", subtitle = "Data from: Abrahamson, W.G. 2019")

```

**Figure 2**: ADD TEXT HERE

```{r}
ggplot(data = palmetto_clean, aes(x = length, y = green_lvs)) +
  geom_point(aes(color = species)) +
  theme_minimal() +
  labs(x = "Length (cm)", y = "Green Leaves (count)", title = "Length and Green Leaves Count for Serenoa repens vs Sabal etonia", subtitle = "Data from: Abrahamson, W.G. 2019")
```

**Figure 3**: ADD TEXT HERE

A section containing 2 - 3 finalized (customized, suitable for a publication) data visualizations (with figure captions) in which you explore differences in height, canopy length, canopy width, and green leaves for the two species. If you prefer, combine the figures into a compound figure using {patchwork} or {cowplot}. Below your data visualizations, add a sentence or two with a takeaway from the plots, e.g., based on these plots, which predictor variables are more likely to help classify species correctly?






## X. Data Citation

Data source: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5