---
title: "Assignment 2: Task 1"
author: "Sophia Leiker"
date: "1/30/2022"
output: 
  html_document: 
    toc: yes
    code_folding: hide
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(patchwork)
library(cowplot)
```


## 1. Introduction

This report will employ binary logistic regression to test feasibility of using variables plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs) to classify whether a palmetto is species Serenoa repens or Sabal etonia.

The analysis will be performed twice, and will use cross validation to compare two models:

- Log odds of plant type using plant height, canopy length, canopy width and green leaves as predictor variable.
- Log odds of plant type using plant height, canopy width and green leaves (i.e., drop canopy length for this model)


**Data Source:** The data used comes from: [Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative](https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5) and metadata, including methods used for data collection can be found here [EDI Data Portal](https://portal.edirepository.org/nis/metadataviewer?packageid=edi.317.1).
The first dataset, palmetto_data, contains survival and growth data across multiple years, habitats and experimental treatments. The second dataset, seedlings_data, follows the fate of marked putative palmetto seedlings in the field to assess survivorship and growth. The final dataset, harvested_palmetto_data, contains size data and estimated dry mass (biomass in grams) of 33 destructively harvested palmetto plants (17 S. repens and 16 S. etonia) of varying sizes and across habitats.



```{r}
# Reading in the data
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  clean_names()

```

***

## X. Data Citation

Data source: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5